\Large{Adversarial Training against Systematic Uncertainty}

### 2016-09-29 tex/paper.tex
save

Systematic uncertainties are parameterized by nuisance parameters in the
statistical analysis of data (see e.g., \citep{bohm2010introduction,cowan2011asymptotic}).  Nuisance parameter often,
though not always, have a suitable prior coming from alternative measurements in
data.  For instance, the average measured energy of an electron in simulation
may be compared with those in data using a relatively pure sample of $Z$ boson
decays to electrons.  The average measured electron energy can not be measured
without uncertainty, and the measured uncertainty can be used in other analyses
to estimate the uncertainty deriving from our imperfect knowledge of the
electron's energy.  A nuisance parameter is assigned to the electron energy
uncertainty, which controls the variation of the electron energy in the
statistical analysis of data.  Nuisance parameters are typically, though not
always, constrained by a suitable prior, e.g. a Gaussian distribution with a
standard deviation equal to the measured uncertainty. \glnote{... and
accordingly, we want to build a classifier independent on the nuisance
parameter, as a way to be independent of the corresponding systematics.}

@@

To account for systematic uncertainties, experimentalists in high energy physics
typically take as fixed a classifier $f$ built from training data for a nominal
value $z_0$ of the nuisance parameter, and then propagate uncertainty
\glnote{add ref} by estimating $p(f(x)|z)$ with a parameterized calibration
procedure. Clearly, this classifier is however not optimal for $z \neq z_0$.
To overcome this issue, the classifier $f$ is sometimes \glnote{add refs} built instead on a mixture
of training data generated from several nominal values $z_0, z_1, \dots$ of the nuisance.
While this certainly improves with respect to classification performance,
there is however no guarantee that the resulting classifier is pivotal, as shown
previously in Section~\ref{sec:toy}.
As an alternative, parameterized
classifiers~\citep{cranmer2015approximating,Baldi:2016fzo} directly take
(nuisance) parameters as additional input variables, hence ultimately providing
the most statistically powerful approach for incorporating the effect of
systematics on the underlying classification task.  As argued in
\citep{Neal:2007zz}, such classifiers can however not be used on real data since
the correct value $z$ of the nuisance often remains unknown. This is typically
not an issue in the context of parameter
inference~\citep{cranmer2015approximating}, where nuisance parameters are
marginalized out, but otherwise often limits the range of their applications. In
practice, parameterized classifiers  are also computationally expensive to build
and evaluate. In particular, calibrating their decision function, i.e.
approximating $p(f(x,z)|z)$ as a continuous function of $z$, remains an open
challenge. By contrast, constraining $f$ to be pivotal yields a classifier which
may not be optimal with respect to $Y|X,Z$, as discussed in
Section~\ref{sec:theory}, but that can otherwise be used in a wider range of
applications, since knowing the correct value $z$ of the nuisance is not
necessary. Similarly, calibration needs to be carried out only once, since  the
dependence on the nuisance is now built-in.

>>>>

Systematic uncertainties are parameterized by nuisance parameters in the
statistical analysis of data (see e.g.,
\citep{bohm2010introduction,cowan2011asymptotic}).  Nuisance parameters often,
though not always, have a suitable prior coming from alternative measurements in
data.  For instance, the average measured energy of an electron in simulation
may be compared with those in data using a relatively pure sample of $Z$ boson
decays to electrons.  The average measured electron energy can not be measured
without uncertainty, and the measured uncertainty can be used in other analyses
to estimate the uncertainty deriving from our imperfect knowledge of the
electron's energy.  A nuisance parameter is assigned to the electron energy
uncertainty, which controls the variation of the electron energy in the
statistical analysis of data.  Nuisance parameters are typically, though not
always, constrained by a suitable prior, e.g. a Gaussian distribution with a
standard deviation equal to the measured uncertainty. \glnote{This paragraph
could be improved, it is not very clear what we want to say...}

@@

To account for systematic uncertainties, experimentalists in high energy physics
typically take as fixed a classifier $f$ built from training data for a nominal
value $z_0$ of the nuisance parameter, and then propagate uncertainty
 by estimating $p(f(x)|z)$ with a parameterized calibration
procedure. Clearly, this classifier is however not optimal for $z \neq z_0$.
To overcome this issue, the classifier $f$ is sometimes built instead on a mixture
of training data generated from several nominal values $z_0, z_1, \dots$ of the nuisance.
While this certainly improves with respect to classification performance,
there is however no guarantee that the resulting classifier is pivotal, as shown
previously in Section~\ref{sec:toy}.
As an alternative, parameterized
classifiers~\citep{cranmer2015approximating,Baldi:2016fzo} directly take
(nuisance) parameters as additional input variables, hence ultimately providing
the most statistically powerful approach for incorporating the effect of
systematics on the underlying classification task.  As argued in
\citep{Neal:2007zz}, such classifiers can however not be used on real data since
the correct value $z$ of the nuisance often remains unknown. This is typically
not an issue in the context of parameter
inference~\citep{cranmer2015approximating}, where nuisance parameters are
marginalized out, but otherwise often limits the range of their applications. In
practice, parameterized classifiers  are also computationally expensive to build
and evaluate. In particular, calibrating their decision function, i.e.
approximating $p(f(x,z)|z)$ as a continuous function of $z$, remains an open
challenge. By contrast, constraining $f$ to be pivotal yields a classifier which
may not be optimal with respect to $Y|X,Z$, as discussed in
Section~\ref{sec:theory}, but that can otherwise be used in a wider range of
applications, since knowing the correct value $z$ of the nuisance is not
necessary. Similarly, calibration needs to be carried out only once, since  the
dependence on the nuisance is now built-in.
